Remote Collection of Respiratory Samples (NACFC 2023)

### 2023-03-29 Jordana Hoppe/nacfc_2023.qmd
Update nacfc_2023.qmd
tidyverse formatting

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(gtsummary)
df = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv",na.strings = "")
surveys = read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv",na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"]="Other"
df$Mutation.2[df$Mutation.2 != "F508del"]="Other"
df$Genotype = paste(df$Mutation.1,df$Mutation.2,sep = "/")
# Age
df$Age = difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB)/365.25
# Combine race columns
df = df %>% rowwise() %>%
  mutate(Race = paste(which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
                                       Race..choice.Unknown.or.not.reported.)== "Checked"),
                      collapse = "+"),
         Race = factor(Race,levels = c("","1","5"),labels=c("Unknown","American Indian or Alaska Native","White")))
# Shorten organism names
colnames(df) = sub("If.culture.was.positive..select.all.organisms.isolated...choice.","",colnames(df))
```

@@

```{r}
#| label: tbl-1
df %>% select(Age,FEV1...predicted,FVC...predicted,
              Sex,Race,Hispanic.or.Latinx,Rural.location.) %>%
  tbl_summary(
    label = list(
      Age ~ "Age at clinic culture collection",
      FEV1...predicted~"ppFEV",FVC...predicted~"ppFVC",
      Hispanic.or.Latinx~"Hispanic or Latinx",
      Rural.location. ~"Rural location?"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ))
```

@@

```{r}
sens_spec = function(gold_standard,other_test,bug,type = "Home Collection"){
  t = table(other_test,gold_standard)
  if(all(dim(t)==c(1,1))){
    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    cat("All samples marked as",unique(gold_standard))
    cat("\n")
  } else {
    sen = sensitivity(data = factor(other_test),reference = factor(gold_standard))
    spec = specificity(data = factor(other_test),reference = factor(gold_standard))

    dimnames(t)=lapply(dimnames(t),function(n){
      n = sub("Checked","Present",n)
      n = sub("Unchecked","Absent",n)
    })
    dimnames(t)$other_test = paste(type,dimnames(t)$other_test)
    dimnames(t)$gold_standard = paste("Clinic",dimnames(t)$gold_standard)

    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    print(knitr::kable((as.data.frame.matrix(t))))
    cat("\n")
    cat("Sensitivity of the",tolower(type),"method was",round(sen,3),"and specificity was",round(spec,3))
    cat("\n")
  }
}
```

@@

```{r results='asis'}
sens_spec(df$Methicillin.sensitive.Staph.aureus.,
          df$Methicillin.sensitive.Staph.aureus..1,
          type = "Home Collection",bug = "MSSA")
```

```{r results='asis'}
sens_spec(df$Methicillin.resistant.Staph.aureus.,
          df$Methicillin.resistant.Staph.aureus..1,
          type = "Home Collection",bug = "MRSA")
```

```{r results='asis'}
sens_spec(df$A..xylosoxidans.,df$A..xylosoxidans..1,
          type = "Home Collection",bug = "A. xylosoxidans")
```

```{r results='asis'}
sens_spec(df$Achromobacter.not.xylosoxidans.,df$Achromobacter.not.xylosoxidans..1,
          type = "Home Collection",bug = "Achromobacter (not xylosoxidans)")
```

```{r results='asis'}
sens_spec(df$Acinetobacter.,df$Acinetobacter..1,
          type = "Home Collection",bug = "Acinetobacter")
```

```{r results='asis'}
sens_spec(df$H..influenzae.,df$H..influenzae..1,
          type = "Home Collection",bug = "H. influenzae")
```

```{r results='asis'}
sens_spec(df$P..aeruginosa.,df$P..aeruginosa..1,
          type = "Home Collection",bug = "P. aeruginosa")
```

```{r results='asis'}
sens_spec(df$P..aeruginosa..mucoid..,df$P..aeruginosa..mucoid...1,
          type = "Home Collection",bug = "P. aeruginosa (mucoid)")
```

```{r results='asis'}
sens_spec(df$Pseudomonas.species..not.aeruginosa.,df$Pseudomonas.species..not.aeruginosa..1,
          type = "Home Collection",bug = "Pseudomonas (not aeruginosa)")
```

```{r results='asis'}
sens_spec(df$B..cepacia.complex.,df$B..cepacia.complex..1,
          type = "Home Collection",bug = "B. cepacia complex")
```

```{r results='asis'}
sens_spec(df$Stenotrophomonas.,df$Stenotrophomonas..1,
          type = "Home Collection",bug = "Stenotrophomonas")
```

```{r results='asis'}
sens_spec(df$Aspergillus.,df$Aspergillus..1,
          type = "Home Collection",bug = "Aspergillus")
```

```{r results='asis'}
sens_spec(df$Penicillium.species.,df$Penicillium.species..1,
          type = "Home Collection",bug = "Penicillium species")
```

```{r results='asis'}
sens_spec(df$Sceosporium.species.,df$Sceosporium.species..1,
          type = "Home Collection",bug = "Sceosporium species")
```

```{r results='asis'}
sens_spec(df$M..avium.complex.,df$M..avium.complex..1,
          type = "Home Collection",bug = "M. avium complex")
```

```{r results='asis'}
sens_spec(df$M..abscessus.,df$M..abscessus..1,
          type = "Home Collection",bug = "M. abscessus")
```

>>>>

```{r}
#| include: false
library(pander)
library(tidyverse)
library(caret)
library(gtsummary)
df <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/clinical.csv", na.strings = "")
surveys <- read.csv("/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Jordana Hoppe/Home Collection/Data_Clean/surveys.csv", na.strings = "")
# Combined genotype
df$Mutation.1[df$Mutation.1 != "F508del"] <- "Other"
df$Mutation.2[df$Mutation.2 != "F508del"] <- "Other"
df$Genotype <- paste(df$Mutation.1, df$Mutation.2, sep = "/")
# Age
df$Age <- difftime(df$Date.of.clinic.culture.collection, df$Patient.DOB) / 365.25
# Combine race columns
df <- df %>%
  rowwise() %>%
  mutate(
    Race = paste(
      which(c_across(Race..choice.American.Indian.or.Alaska.Native.:
      Race..choice.Unknown.or.not.reported.) == "Checked"),
      collapse = "+"
    ),
    Race = factor(Race, levels = c("", "1", "5"), labels = c("Unknown", "American Indian or Alaska Native", "White"))
  )
# Shorten organism names
colnames(df) <- sub("If.culture.was.positive..select.all.organisms.isolated...choice.", "", colnames(df))
```

@@

```{r}
#| label: tbl-1
df %>%
  select(
    Age, FEV1...predicted, FVC...predicted,
    Sex, Race, Hispanic.or.Latinx, Rural.location.
  ) %>%
  tbl_summary(
    label = list(
      Age ~ "Age at clinic culture collection",
      FEV1...predicted ~ "ppFEV", FVC...predicted ~ "ppFVC",
      Hispanic.or.Latinx ~ "Hispanic or Latinx",
      Rural.location. ~ "Rural location?"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    )
  )
```

@@

```{r}
sens_spec <- function(gold_standard, other_test, bug, type = "Home Collection") {
  t <- table(other_test, gold_standard)
  if (all(dim(t) == c(1, 1))) {
    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    cat("All samples marked as", unique(gold_standard))
    cat("\n")
  } else {
    sen <- sensitivity(data = factor(other_test), reference = factor(gold_standard))
    spec <- specificity(data = factor(other_test), reference = factor(gold_standard))

    dimnames(t) <- lapply(dimnames(t), function(n) {
      n <- sub("Checked", "Present", n)
      n <- sub("Unchecked", "Absent", n)
    })
    dimnames(t)$other_test <- paste(type, dimnames(t)$other_test)
    dimnames(t)$gold_standard <- paste("Clinic", dimnames(t)$gold_standard)

    cat("\n")
    cat("###", bug)
    cat("\n")
    cat("#### Contingency table")
    cat("\n")
    print(knitr::kable((as.data.frame.matrix(t))))
    cat("\n")
    cat("Sensitivity of the", tolower(type), "method was", round(sen, 3), "and specificity was", round(spec, 3))
    cat("\n")
  }
}
```

@@

```{r results='asis'}
sens_spec(df$Methicillin.sensitive.Staph.aureus.,
  df$Methicillin.sensitive.Staph.aureus..1,
  type = "Home Collection", bug = "MSSA"
)
```

```{r results='asis'}
sens_spec(df$Methicillin.resistant.Staph.aureus.,
  df$Methicillin.resistant.Staph.aureus..1,
  type = "Home Collection", bug = "MRSA"
)
```

```{r results='asis'}
sens_spec(df$A..xylosoxidans., df$A..xylosoxidans..1,
  type = "Home Collection", bug = "A. xylosoxidans"
)
```

```{r results='asis'}
sens_spec(df$Achromobacter.not.xylosoxidans., df$Achromobacter.not.xylosoxidans..1,
  type = "Home Collection", bug = "Achromobacter (not xylosoxidans)"
)
```

```{r results='asis'}
sens_spec(df$Acinetobacter., df$Acinetobacter..1,
  type = "Home Collection", bug = "Acinetobacter"
)
```

```{r results='asis'}
sens_spec(df$H..influenzae., df$H..influenzae..1,
  type = "Home Collection", bug = "H. influenzae"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa., df$P..aeruginosa..1,
  type = "Home Collection", bug = "P. aeruginosa"
)
```

```{r results='asis'}
sens_spec(df$P..aeruginosa..mucoid.., df$P..aeruginosa..mucoid...1,
  type = "Home Collection", bug = "P. aeruginosa (mucoid)"
)
```

```{r results='asis'}
sens_spec(df$Pseudomonas.species..not.aeruginosa., df$Pseudomonas.species..not.aeruginosa..1,
  type = "Home Collection", bug = "Pseudomonas (not aeruginosa)"
)
```

```{r results='asis'}
sens_spec(df$B..cepacia.complex., df$B..cepacia.complex..1,
  type = "Home Collection", bug = "B. cepacia complex"
)
```

```{r results='asis'}
sens_spec(df$Stenotrophomonas., df$Stenotrophomonas..1,
  type = "Home Collection", bug = "Stenotrophomonas"
)
```

```{r results='asis'}
sens_spec(df$Aspergillus., df$Aspergillus..1,
  type = "Home Collection", bug = "Aspergillus"
)
```

```{r results='asis'}
sens_spec(df$Penicillium.species., df$Penicillium.species..1,
  type = "Home Collection", bug = "Penicillium species"
)
```

```{r results='asis'}
sens_spec(df$Sceosporium.species., df$Sceosporium.species..1,
  type = "Home Collection", bug = "Sceosporium species"
)
```

```{r results='asis'}
sens_spec(df$M..avium.complex., df$M..avium.complex..1,
  type = "Home Collection", bug = "M. avium complex"
)
```

```{r results='asis'}
sens_spec(df$M..abscessus., df$M..abscessus..1,
  type = "Home Collection", bug = "M. abscessus"
)
```
